Solving the maximum independant set problem with tensor networks

### 2021-05-25 paper/paper.tex
update paper

\iclrfinalcopy
\begin{document}
\title{Solving the maximum independant set problem with Tropical tensor networks}

\author{Jin-Guo Liu\\
Harvard University\\
\texttt{jinguoliu@harvard.edu}\\
%\AND
%Sheng-Tao Wang\\
%QuEra computing Inc.
}
\maketitle

\begin{abstract}
	Solving the maximum independent set size problem: the maximum independent set size, the degeneracy, the optimal configuration and the equivalence between different graphs.
\end{abstract}

\section{Tools}
\begin{description}
	\item[OMEinsum] a package for einsum \\ \href{https://github.com/under-Peter/OMEinsum.jl}{https://github.com/under-Peter/OMEinsum.jl}
	\item[TropicalGEMM] a package for efficient tropical matrix multiplication (compatible with OMEinsum) \\ \href{https://github.com/TensorBFS/TropicalGEMM.jl}{https://github.com/TensorBFS/TropicalGEMM.jl}
	\item[TropicalNumbers] a package providing tropical number types and tropical algebra, one o the dependency of TropicalGEMM \\ \href{https://github.com/TensorBFS/TropicalNumbers.jl}{https://github.com/TensorBFS/TropicalNumbers.jl}
\end{description}

@@

If put a vector of size $2$ at each vertex
\begin{equation}
    \left(\begin{matrix}
        1 \\
        x
    \end{matrix}\right),
\end{equation}

and a matrix of size $2 \times 2$ at each edge
\begin{equation}
    \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right).
\end{equation}

@@

Here $x$ is a variable. Then we contract this graph with einsum \texttt{a,b,c,d,ab,ad,bc,bd,cd,ce->}.
The result is a scalar function of $x$, it corresponds to the independance polynomial~\cite{}.
\begin{equation}
I(G, x) = \sum_{k=1}^{\alpha(G)} a_k x^k,
\end{equation}
where $a_k$ is the the degeneracy for independant set size $k$.

>>>>

\iclrfinalcopy
\begin{document}
\title{Solving the maximum independant set problem with tensor networks}

\author{Jin-Guo Liu\\
Harvard University\\
\texttt{jinguoliu@g.harvard.edu}\\
\AND
Xun Gao\\
Harvard University\\
\texttt{xungao@g.harvard.edu}\\
%\AND
%Sheng-Tao Wang\\
%QuEra computing Inc.
}
\maketitle

\begin{abstract}
	Solving the maximum independent set size problem: the maximum independent set size, the independance polynomial and the optimal configuration.
\end{abstract}

\section{Technical guide}
\begin{description}
	\item[OMEinsum] a package for einsum,
	\item[OMEinsumContractionOrders] a package for finding the optimal contraction order for einsum \\ \href{https://github.com/Happy-Diode/OMEinsumContractionOrders.jl}{https://github.com/Happy-Diode/OMEinsumContractionOrders.jl},
	\item[TropicalGEMM] a package for efficient tropical matrix multiplication (compatible with OMEinsum),
	\item[TropicalNumbers] a package providing tropical number types and tropical algebra, one o the dependency of TropicalGEMM,
	\item[LightGraphs] a package providing graph utilities, like random regular graph generator,
	\item[Polynomials] a package providing polynomial algebra and polynomial fitting,
	\item[Mods and Primes] packages providing finite field algebra and prime number generators.
\end{description}

One can install these packages by opening a julia REPL, type \colorbox{lightgray}{\texttt{]}} to enter the \texttt{pkg>} mode and type, e.g.
\begin{lstlisting}
pkg> add OMEinsum
\end{lstlisting}

An example of computing the maximum independant set size
\begin{lstlisting}
julia> using OMEinsum, LightGraphs, TropicalNumbers

julia> n, k = 6, 3
(6, 3)

julia> g = LightGraphs.random_regular_graph(n, k)
{6, 9} undirected simple Int64 graph

julia> ixs = [minmax(e.src,e.dst) for e in LightGraphs.edges(g)]
9-element Vector{Tuple{Int64, Int64}}:
 (1, 2)
 (1, 4)
 (1, 5)
 (2, 3)
 (2, 6)
 (3, 5)
 (3, 6)
 (4, 5)
 (4, 6)

julia> code = EinCode((ixs..., [(i,) for i in LightGraphs.vertices(g)]...), ())
1∘2, 1∘4, 1∘5, 2∘3, 2∘6, 3∘5, 3∘6, 4∘5, 4∘6, 1, 2, 3, 4, 5, 6 ->

julia> optimized_code = optimize_greedy(code, Dict([i=>2 for i=1:6]))
4∘6, 4∘6 -> 
├─ 4∘6, 6 -> 4∘6
│  ├─ 6
│  └─ 4∘6
└─ 2∘5∘4, 5∘2∘6 -> 4∘6
   ├─ 2∘5∘6, 6∘2 -> 5∘2∘6
   │  ├─ 2∘6, 2 -> 6∘2
   │  │  ├─ 2
   │  │  └─ 2∘6
   │  └─ 2∘3, 5∘6∘3 -> 2∘5∘6
   │     ├─ 3∘5, 6∘3 -> 5∘6∘3
   │     │  ├─ 3∘6, 3 -> 6∘3
   │     │  │  ⋮
   │     │  │  
   │     │  └─ 3∘5, 5 -> 3∘5
   │     │     ⋮
   │     │     
   │     └─ 2∘3
   └─ 2∘4∘1, 1∘4∘5 -> 2∘5∘4
      ├─ 1∘5, 5∘4 -> 1∘4∘5
      │  ├─ 4∘5, 4 -> 5∘4
      │  │  ├─ 4
      │  │  └─ 4∘5
      │  └─ 1∘5
      └─ 2∘1, 1∘4 -> 2∘4∘1
         ├─ 1∘4
         └─ 1∘2, 1 -> 2∘1
            ├─ 1
            └─ 1∘2

julia> function mis_contract(code::OMEinsum.NestedEinsum, x::T) where T
           tensors = map(OMEinsum.getixs(Iterators.flatten(code))) do ix
               @assert length(ix) == 1 || length(ix) == 2
                           length(ix) == 1 ? [one(T), x] : [one(T) one(T); one(T) zero(T)]
           end
                 code(tensors...)
       end
mis_contract (generic function with 1 method)

julia> result = mis_contract(optimized_code, Tropical(1.0))
0-dimensional Array{TropicalF64, 0}:
2.0ₜ
\end{lstlisting}

For larger graph, we hight recommend using two additional packages,
\texttt{TropicalGEMM} for BLAS speed tropical matrix multiplication and
\texttt{OMEinsumContractionOrders} for hyper optimized contraction order (the KaHyPar approach) ~\cite{Gray2021,Pan2021}.

@@

If put a vector of size $2$ at each vertex
\begin{equation}
    \left(\begin{matrix}
        1 \\
        x
    \end{matrix}\right),
\end{equation}
were $x$ is a variable, and a matrix of size $2 \times 2$ at each edge
\begin{equation}
    \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right).
\end{equation}

@@

Then we contract this graph with einsum \texttt{a,b,c,d,ab,ad,bc,bd,cd,ce->}.
The result is a scalar function of $x$, it corresponds to the independance polynomial~\cite{}.
\begin{equation}
I(G, x) = \sum_{k=1}^{\alpha(G)} a_k x^k,
\end{equation}
where $a_k$ is the the degeneracy for independant set size $k$.
